---
title: "hw6"
author: "Lulu Zhang"
date: "11/21/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE)

library(tidyverse)
library(modelr)
library(mgcv)
library(viridis)
```
# Problem 1

I will load and tidy the dataset `birthweight` for regression analysis. I converted `frace`, `mrace`, `babysex`, 
and `malform` to factors; confirmed no missing variables; stnadardized variable to the metric system
so i converted `delwt`, `ppwt`, and `wtgain`, from pounds to grams, and `mheight` from inches to cm. 

```{r}
# load data
bwt = read.csv("./birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
   frace = as.factor(frace),
   frace = recode(frace,
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other", 
    "9" = "Unknown"),
   mrace = as.factor(mrace),
   mrace = recode(mrace, 
    "1" = "White", 
    "2" = "Black", 
    "3" = "Asian", 
    "4" = "Puerto Rican", 
    "8" = "Other"),
   babysex = as.factor(babysex),
   babysex = recode(babysex, 
     "1" = "male", 
    "2" = "female"),
   malform = as.factor(malform),
   malform = recode(malform,
    "0" = "absent", 
    "1" = "present"),
   delwt = 453.592*delwt, 
   ppwt = 453.592*ppwt, 
   wtgain = 453.592*wtgain, 
   mheight = 2.54*mheight)
  
anyNA(bwt)
# no missing values
```

Proposing a model for birthweight. I hypothesize `gaweeks`: gestational age in weeks, `pnumlbw`: previous number of low birth weight babies,
`pnumgsa`:  number of prior small for gestational age babies, `ppbmi`:mother’s pre-pregnancy BMI, `smoken`: average number of cigarettes
smoked per day during pregnancy, and `ppwt`: mother's pre-pregnancy weight are a few of the factors that influence baby's birthweight (outcome). 


```{r include=FALSE}
bwt %>% 
  mutate(
    mrace = fct_infreq(mrace)
  ) %>% 
  select(gaweeks, pnumlbw, pnumsga, ppbmi, smoken, ppwt)
fit = lm(bwt ~ gaweeks + pnumlbw + pnumsga + smoken + ppwt, data = bwt)
fit %>% 
  broom::tidy()
```
Compare your model to two others:

One using length at birth and gestational age as predictors (main effects only)
One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

```{r}
# main effects model
fit_main = lm(bwt ~ blength + gaweeks, data = bwt)
fit_main %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  rename(
    "Term" = "term",
    "Estimate" = "estimate",
    "p-value" = "p.value") %>% 
  knitr::kable(digits = 4)

# interactions model
fit_interact = lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex +  bhead * blength * babysex, data = bwt)
fit_interact %>%
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  rename(
    "Term" = "term",
    "Estimate" = "estimate",
    "p-value" = "p.value") %>% 
  knitr::kable(digits = 4)
```


# Problem 2

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

